Mitigating Distributional Biases in Contrastive Learning

### 2023-05-03 paper/main.tex
citations check

Contrastive learning has proven to be effective for eliciting visual representations: SimCLR \citep{Chen2020SimCLR} is a framework that considerably outperforms previous self-supervised and semi-supervised learning methods. Its authors developed NT-Xent loss (normalized temperature-scaled cross-entropy loss), that uses cosine similarity instead of scalar products.

During training, the lack of true labels often necessitates the random selection of negative instances $\textbf{x}^-$ from the training data. However, this approach results in sampling bias, where $\textbf{x}^-$ may actually be similar to $\textbf{x}$. This bias leads to a significant drop in performance \citep{hong2021unbiased}.

>>>>

Contrastive learning has proven to be effective for eliciting visual representations: SimCLR \citep{chen2020simple} is a framework that considerably outperforms previous self-supervised and semi-supervised learning methods. Its authors developed NT-Xent loss (normalized temperature-scaled cross-entropy loss), that uses cosine similarity instead of scalar products.

During training, the lack of true labels often necessitates the random selection of negative instances $\textbf{x}^-$ from the training data. However, this approach results in sampling bias, where $\textbf{x}^-$ may actually be similar to $\textbf{x}$. This bias leads to a significant drop in performance \citep{NEURIPS2021_de8aa43e}.
